A Short Demo Article: Regression Models for Count Data in \proglang{R}

### 2022-06-21 5_artigos/journal-of-statistical-software/template-revista/article-knitr.tex
corrige texto, atualiza artigos

%% recommended packages
%\usepackage{orcidlink,thumbpdf,lmodern}
\usepackage{thumbpdf,lmodern}

@@

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://www.zeileis.org/}
}

\begin{document}
\input{article-concordance}

@@

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
\begin{Schunk}
\begin{Sinput}
R> data("quine", package = "MASS")
\end{Sinput}
\end{Schunk}
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

@@

\begin{figure}[t!]
\centering
\includegraphics{article-visualization}
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
\begin{Schunk}
\begin{Sinput}
R> m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
+    family = poisson)
\end{Sinput}
\end{Schunk}
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
\begin{Schunk}
\begin{Sinput}
R> library("MASS")
R> m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
\end{Sinput}
\end{Schunk}
%
In a comparison with the BIC the latter model is clearly preferred.
%
\begin{Schunk}
\begin{Sinput}
R> BIC(m_pois, m_nbin)
\end{Sinput}
\begin{Soutput}
       df      BIC
m_pois 18 2046.851
m_nbin 19 1157.235
\end{Soutput}
\end{Schunk}
%
Hence, the full summary of that model is shown below.
%
\begin{Schunk}
\begin{Sinput}
R> summary(m_nbin)
\end{Sinput}
\begin{Soutput}
Call:
glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
    init.theta = 1.60364105, link = log)

Deviance Residuals: 
    Min       1Q   Median       3Q      Max  
-3.0857  -0.8306  -0.2620   0.4282   2.0898

Coefficients: (1 not defined because of singularities)
            Estimate Std. Error z value Pr(>|z|)    
(Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
EthN        -0.24591    0.39135  -0.628  0.52977    
SexM        -0.77181    0.38021  -2.030  0.04236 *  
AgeF1       -0.02546    0.41615  -0.061  0.95121    
AgeF2       -0.54884    0.54393  -1.009  0.31296    
AgeF3       -0.25735    0.40558  -0.635  0.52574    
LrnSL        0.38919    0.48421   0.804  0.42153    
EthN:SexM    0.36240    0.29430   1.231  0.21818    
EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
AgeF3:LrnSL       NA         NA      NA       NA    
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

(Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)

    Null deviance: 235.23  on 145  degrees of freedom
Residual deviance: 167.53  on 128  degrees of freedom
AIC: 1100.5

Number of Fisher Scoring iterations: 1

              Theta:  1.604 
          Std. Err.:  0.214

 2 x log-likelihood:  -1062.546 
\end{Soutput}
\end{Schunk}

>>>>

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

@@

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://www.zeileis.org/}
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{data}\hlstd{(}\hlstr{"quine"}\hlstd{,} \hlkwc{package} \hlstd{=} \hlstr{"MASS"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

@@

\begin{figure}[t!]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/visualization-1} 
\end{knitrout}
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{m_pois} \hlkwb{<-} \hlkwd{glm}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine,}
  \hlkwc{family} \hlstd{= poisson)}
\end{alltt}
\end{kframe}
\end{knitrout}
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{library}\hlstd{(}\hlstr{"MASS"}\hlstd{)}
\hlstd{m_nbin} \hlkwb{<-} \hlkwd{glm.nb}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine)}
\end{alltt}
\end{kframe}
\end{knitrout}
%
In a comparison with the BIC the latter model is clearly preferred.
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{BIC}\hlstd{(m_pois, m_nbin)}
\end{alltt}
\begin{verbatim}
##        df      BIC
## m_pois 18 2046.851
## m_nbin 19 1157.235
\end{verbatim}
\end{kframe}
\end{knitrout}
%
Hence, the full summary of that model is shown below.
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{summary}\hlstd{(m_nbin)}
\end{alltt}
\begin{verbatim}
## 
## Call:
## glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
##     init.theta = 1.60364105, link = log)
## 
## Deviance Residuals: 
##     Min       1Q   Median       3Q      Max  
## -3.0857  -0.8306  -0.2620   0.4282   2.0898  
## 
## Coefficients: (1 not defined because of singularities)
##             Estimate Std. Error z value Pr(>|z|)    
## (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
## EthN        -0.24591    0.39135  -0.628  0.52977    
## SexM        -0.77181    0.38021  -2.030  0.04236 *  
## AgeF1       -0.02546    0.41615  -0.061  0.95121    
## AgeF2       -0.54884    0.54393  -1.009  0.31296    
## AgeF3       -0.25735    0.40558  -0.635  0.52574    
## LrnSL        0.38919    0.48421   0.804  0.42153    
## EthN:SexM    0.36240    0.29430   1.231  0.21818    
## EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
## EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
## EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
## EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
## SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
## SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
## SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
## SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
## AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
## AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
## AgeF3:LrnSL       NA         NA      NA       NA    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
## 
##     Null deviance: 235.23  on 145  degrees of freedom
## Residual deviance: 167.53  on 128  degrees of freedom
## AIC: 1100.5
## 
## Number of Fisher Scoring iterations: 1
## 
## 
##               Theta:  1.604 
##           Std. Err.:  0.214 
## 
##  2 x log-likelihood:  -1062.546
\end{verbatim}
\end{kframe}
\end{knitrout}

### 2022-06-21 5_artigos/journal-of-statistical-software/template-revista/article.tex
corrige texto, atualiza artigos

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
@

@@

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://www.zeileis.org/}
}

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
<<data>>=
data("quine", package = "MASS")
@
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

@@

\begin{figure}[t!]
\centering
<<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
par(mar = c(4, 4, 1, 1))
plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
axis(2)
axis(1, at = 0:16 * 5, labels = FALSE)
axis(1, at = 0:8 * 10)
box()
@
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
<<poisson>>=
m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
  family = poisson)
@
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
<<negbin>>=
library("MASS")
m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
@
%
In a comparison with the BIC the latter model is clearly preferred.
%
<<comparison>>=
BIC(m_pois, m_nbin)
@
%
Hence, the full summary of that model is shown below.
%
<<summary>>=
summary(m_nbin)
@

>>>>

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}

@@

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://www.zeileis.org/}
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{data}\hlstd{(}\hlstr{"quine"}\hlstd{,} \hlkwc{package} \hlstd{=} \hlstr{"MASS"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

@@

\begin{figure}[t!]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/visualization-1} 
\end{knitrout}
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{m_pois} \hlkwb{<-} \hlkwd{glm}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine,}
  \hlkwc{family} \hlstd{= poisson)}
\end{alltt}
\end{kframe}
\end{knitrout}
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{library}\hlstd{(}\hlstr{"MASS"}\hlstd{)}
\hlstd{m_nbin} \hlkwb{<-} \hlkwd{glm.nb}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine)}
\end{alltt}
\end{kframe}
\end{knitrout}
%
In a comparison with the BIC the latter model is clearly preferred.
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{BIC}\hlstd{(m_pois, m_nbin)}
\end{alltt}
\begin{verbatim}
##        df      BIC
## m_pois 18 2046.851
## m_nbin 19 1157.235
\end{verbatim}
\end{kframe}
\end{knitrout}
%
Hence, the full summary of that model is shown below.
%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{summary}\hlstd{(m_nbin)}
\end{alltt}
\begin{verbatim}
## 
## Call:
## glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
##     init.theta = 1.60364105, link = log)
## 
## Deviance Residuals: 
##     Min       1Q   Median       3Q      Max  
## -3.0857  -0.8306  -0.2620   0.4282   2.0898  
## 
## Coefficients: (1 not defined because of singularities)
##             Estimate Std. Error z value Pr(>|z|)    
## (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
## EthN        -0.24591    0.39135  -0.628  0.52977    
## SexM        -0.77181    0.38021  -2.030  0.04236 *  
## AgeF1       -0.02546    0.41615  -0.061  0.95121    
## AgeF2       -0.54884    0.54393  -1.009  0.31296    
## AgeF3       -0.25735    0.40558  -0.635  0.52574    
## LrnSL        0.38919    0.48421   0.804  0.42153    
## EthN:SexM    0.36240    0.29430   1.231  0.21818    
## EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
## EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
## EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
## EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
## SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
## SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
## SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
## SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
## AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
## AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
## AgeF3:LrnSL       NA         NA      NA       NA    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
## 
##     Null deviance: 235.23  on 145  degrees of freedom
## Residual deviance: 167.53  on 128  degrees of freedom
## AIC: 1100.5
## 
## Number of Fisher Scoring iterations: 1
## 
## 
##               Theta:  1.604 
##           Std. Err.:  0.214 
## 
##  2 x log-likelihood:  -1062.546
\end{verbatim}
\end{kframe}
\end{knitrout}
